---
title: "Cardiovascular Disease Prediction Visualizations (By: Frenkli Axha)"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r, include=FALSE}
library(tidyverse)
library(dplyr)
library(data.table)
library(magrittr)
library(ggplot2)
library(ggthemes)
library(hrbrthemes)
df <-read.csv("cardio_train.csv", sep=";")
df$age<- df$age/365
```

# Summary of data set:

*  descriptive stats for all variables
```{r}
#Creating a table with the basic information from the data frame: the minimum value, maximum, mean,
summary(object = df)
```

* structure of our dataset:
```{r}
str(df)
```

* variables:
```{r}
colnames(df)
```

* number of missing values 
```{r}

#In this stage, we want to achieve a clean data set, where there are no duplicates, or missing variables.
#Secondly, we have to extrct all the important variables that we need for the patients such as: Age, gender and height.

number_of_columns <- ncol(df) #here the number of columns/variables is returned
number_of_lines <- nrow(df)   #here the number of rows/observations is returned

#Check whether there are any missing values in the dataframe --> NA
missing_values <- is.na(df)

#Number of missing variables if any
sum(missing_values)
```

* number of unique IDs
```{r}
#Since there are no missing values we can continue with fixing the presentation of the gender variable, by splitting it in two categories for males and females:

#Checking for duplicated IDs:
length(unique(df$id))

```

* range of all variables:
```{r}
apply(df, 2, range)

```

* mean of all variables (as seen also on the summary):
```{r}
colMeans(df)
```

* variance of all variables:
```{r}
apply(df, 2, var)
```

## Boxplots of Age, Weight, Height, Cardio by Age and Active by Age:
* Boxplot Age:
```{r}
boxplot(df$age, main = "Age boxplot", ylab = "Age")
```


* Boxplot height:
```{r}
boxplot(df$height, main = "Height boxplot", ylab = "Height")
```

* Boxplot weight:
```{r}
boxplot(df$weight, main = "Boxplot weight", ylab = "Weight")
```

* Boxplot Cardio:
Boxplots to compare the age and body mass index for the cardio and non-cardio patients.
```{r}
boxplot(df$age~df$cardio, main = "Cardio distriburions per age (in days)", ylab = "Age", xlab = "Cardio")+ theme_clean()
```

* Boxplot Active:
Boxplots to compare the age and body mass index for the active and non-active patients.
```{r}
boxplot(df$age~df$active, main = "Active distriburions per age", ylab = "Age", xlab = "Active")+ theme_clean()
```

## Gender distrbution: Count of Males and Females:

```{r}
ggplot(df, aes(fill=as.factor(gender))) + geom_bar(aes(x = as.factor(gender)))+ scale_fill_hue(c = 40)  + ylab ("Count") + xlab("Gender")+ labs(title="Gender distribution")  +  theme(legend.position="right",axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+ scale_fill_discrete(name = "Gender", labels = c("F", "M"))+ theme_clean()

```

## Age distribution:

```{r}
ggplot(df, aes(x=age)) + 
  geom_histogram(fill = "#FF6A6A") + theme_clean()   + ylab ("Count") + xlab("Age")+ labs(title="Age distribution")
```

## Cholesterol distribution between men and women: 

```{r}
ggplot(df)+geom_bar(aes(x = cholesterol, fill = as.factor(gender)))+xlab("Cholesterol Level") + ylab ("Count") + labs(title="Cholesterol distribution")  +facet_grid(facets = df$gender) + theme_clean()+ scale_fill_discrete(name = "Gender", labels = as.factor(c("F", "M"))) 
```
## Distribution of smoking between men and women:

```{r}
#TODO check gender labels
ggplot(df, aes(fill = as.factor(gender))) + 
  geom_bar(aes(x = as.factor(smoke))) +
  coord_flip()+ scale_fill_hue(c = 40)  + ylab ("Count") + xlab ("Smoke")+ labs(title="Smoke by gender distribution")  +  theme(legend.position="right")+ scale_fill_discrete(name = "Gender", labels = c("F", "M"))+theme_clean()
```
# Explorative statistics

## Activeness by age groups:
```{r}

agebreaks <- c(0,1,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,500)
agelabels <- c("0-1","1-4","5-9","10-14","15-19","20-24","25-29","30-34",
               "35-39","40-44","45-49","50-54","55-59","60-64","65-69",
               "70-74","75-79","80-84","85+")

setDT(df)[ , agegroups := cut(age, 
                                breaks = agebreaks, 
                                right = FALSE, 
                                labels = agelabels)]
plot<- df%>%
  group_by(active)%>%
  count(agegroups)
ggplot(plot, aes(y =n, x = as.factor(active)))+ geom_bar(stat='identity', aes(fill = as.factor(active)))+ facet_wrap(facets = plot$agegroups, scales="free_y")+xlab("Active") + ylab ("Count") + labs(title="Active by age")+ scale_fill_discrete(name = "Active", labels = as.factor(c("Not active", "Active"))) 
```

## Glucose by age groups:
```{r}
plot<- df%>%
    group_by(gluc)%>%
    count(agegroups)

ggplot(plot, aes(y =n, x = as.factor(gluc)))+ geom_bar(stat='identity', aes(fill = as.factor(gluc)))+ facet_wrap(facets = plot$agegroups, scales="free_y")+xlab("Glucose levels") + ylab ("Count") + labs(title="Glucose by age groups")+ scale_fill_discrete(name = "gluc", labels = as.factor(c("Normal", "Above normal", "Way above normal")))
```

## Cholesterol by age groups:
```{r}
plot2<- df%>%
    group_by(cholesterol)%>%
    count(agegroups) 
ggplot(plot2, aes(y =n, x = as.factor(cholesterol)))+ geom_bar(stat='identity', aes(fill = as.factor(cholesterol)))+ facet_wrap(facets = plot$agegroups, scales="free_y")+xlab("Cholesterol levels") + ylab ("Count") + labs(title="Cholesterol by age groups")+ scale_fill_discrete(name = "Cholesterol", labels = as.factor(c("Normal", "Above normal", "Way above normal"))) 
```


